---
title: "Predictive Modelling - Part2"
author: "Vikrant Vaidya"
date: "8/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r include=FALSE}
#IMPORTING LIBRARIES

library(mosaic)
library(quantmod)
library(foreach)
library(dplyr)
library(gridExtra)

```
VISUAL STROY TELLING - FLIGHTS AT ABIA

We are going to take a look at the ABIA data from the eyes and interest of a frequent flyer staying in Austin.
As a frequent flyer I would want to know -

1) What percentage of flights in and out of Austin are delayed ?
2) What is the average amount of times by which flights in and out of Austin get delayed?
3) What are the Origin(flights to Austin) and Destination(flights from Austin) with the most cancellation blues?
4) What causes these cancellations ?
5) What are the airline carriers one should expect most delays with?
6) Which are the best Origin/Dest to expect pre-time arrival/departures to/from Austin ?
7) Which airports do I expect to spend the maximum time Taxing In and Out?
8) What are the times of the day I should expect the maximum arrival and departure delays in/out of Austin?
9) Are these trends better/worse on some days of the week? Whether I could beat this?

But first, PIPING GROUPING AND SUMMARISING the data.
```{r echo=FALSE,warning=FALSE}

abia = read.csv('https://raw.githubusercontent.com/jgscott/STA380/master/data/ABIA.csv')

#Creating Flag Columns
abia <- abia %>% 
  mutate(arr_delay_flag = ifelse(abia$ArrDelay>0,1,0))

abia <- abia %>% 
  mutate(dep_delay_flag = ifelse(abia$DepDelay>0,1,0))

abia$Period_dep[(abia$DepTime >= 500) & (abia$DepTime < 1200)] <- "MORNING"
abia$Period_dep[(abia$DepTime >= 1200) & (abia$DepTime < 1800)] <- "AFTERNOON"
abia$Period_dep[(abia$DepTime >= 1800) | (abia$DepTime < 500)] <- "EVENING"

abia$Period_arr[(abia$ArrTime >= 500) & (abia$ArrTime < 1200)] <- "MORNING"
abia$Period_arr[(abia$ArrTime >= 1200) & (abia$ArrTime < 1800)] <- "AFTERNOON"
abia$Period_arr[(abia$ArrTime >= 1800) | (abia$ArrTime < 500)] <- "EVENING"

#% of DEPARTURES
#Grouping Delays by levels and Plotting
#BY MONTH
depdel_month = abia[(!is.na(abia$dep_delay_flag) & (abia$Origin=="AUS")),] %>%
  group_by(Month) %>%
  summarise(depdel.prt = (sum(dep_delay_flag)/count(Month))*100)

#BY DAY OF MONTH
depdel_dom = abia[(!is.na(abia$dep_delay_flag) & (abia$Origin=="AUS")),] %>%
  group_by(DayofMonth) %>%
  summarize(depdel.prt=(sum(dep_delay_flag)/count(DayofMonth))*100)


#BY DAY OF WEEK
depdel_dow = abia[(!is.na(abia$dep_delay_flag) & (abia$Origin=="AUS")),] %>%
  group_by(DayOfWeek) %>%
  summarize(depdel.prt=(sum(dep_delay_flag)/count(DayOfWeek))*100)


#% of ARRIVALS
#Grouping Delays by levels and Plotting
#BY MONTH
arrdel_month = abia[(!is.na(abia$arr_delay_flag) & (abia$Dest=="AUS")),] %>%
  group_by(Month) %>%
  summarize(arrdel.prt=(sum(arr_delay_flag)/count(Month))*100)

#BY DAY OF MONTH
arrdel_dom = abia[(!is.na(abia$arr_delay_flag) & (abia$Dest=="AUS")),] %>%
  group_by(DayofMonth) %>%
  summarize(arrdel.prt=(sum(arr_delay_flag)/count(DayofMonth))*100)

#BY DAY OF WEEK
arrdel_dow = abia[(!is.na(abia$arr_delay_flag) & (abia$Dest=="AUS")),] %>%
  group_by(DayOfWeek) %>%
  summarize(arrdel.prt=(sum(arr_delay_flag)/count(DayOfWeek))*100)



#2 - Plotting Avg Dep and Arrival Delays by Month/DayofMonth/DayofWeek
#Can get rid of flag use
#DEP DELAYS
mean_depdel_month = abia[((abia$dep_delay_flag==1)& (abia$Origin=="AUS")),] %>%
  group_by(Month) %>%
  summarise(mean.del = mean(DepDelay))

mean_depdel_dom = abia[((abia$dep_delay_flag==1)& (abia$Origin=="AUS")),] %>%
  group_by(DayofMonth) %>%
  summarise(mean.del = mean(DepDelay))

mean_depdel_dow = abia[((abia$dep_delay_flag==1)& (abia$Origin=="AUS")),] %>%
  group_by(DayOfWeek) %>%
  summarise(mean.del = mean(DepDelay))

#ARR DELAYS
mean_arrdel_month = abia[((abia$arr_delay_flag==1) & (abia$Dest=="AUS")),] %>%
  group_by(Month) %>%
  summarise(mean.del = mean(ArrDelay))

mean_arrdel_dom = abia[((abia$arr_delay_flag==1) & (abia$Dest=="AUS")),] %>%
  group_by(DayofMonth) %>%
  summarise(mean.del = mean(ArrDelay))

mean_arrdel_dow = abia[((abia$arr_delay_flag==1) & (abia$Dest=="AUS")),] %>%
  group_by(DayOfWeek) %>%
  summarise(mean.del = mean(ArrDelay))


#3 Cancelled Flights by Origin and Destination/Cancellation Code
orig_cancel = abia[((abia$Cancelled==1) & (abia$Dest=="AUS")),] %>%
  group_by(Origin,CancellationCode) %>%
  summarise(tot.cancel = sum(Cancelled))

dest_cancel = abia[((abia$Cancelled==1) & (abia$Origin=="AUS")),] %>%
  group_by(Dest,CancellationCode) %>%
  summarise(tot.cancel = sum(Cancelled))


#4 Unique Carrier with Carrier Delay
carrier_del = abia[abia$CarrierDelay>=0,] %>%
  group_by(UniqueCarrier) %>%
  summarise(mean.del = mean(CarrierDelay))

#5 Pre-time departures to Austin And Pre-time arrivals from Austin
pre_dep = abia[((!(abia$Origin=="AUS")) & (abia$DepDelay<0) & (!is.na(abia$DepDelay))),] %>%
  group_by(Origin) %>%
  summarise(cnt.dep = n() )

pre_arr = abia[((!(abia$Dest=="AUS")) & (abia$ArrDelay<0) & (!is.na(abia$ArrDelay))),] %>%
  group_by(Dest) %>%
  summarise(cnt.arr = n() )

#6 Mean Taxi-Out/In times by airport
taxi_out = abia[!is.na(abia$TaxiOut),] %>%
  group_by(Origin) %>%
  summarise(mean = mean(TaxiOut))

taxi_in = abia[!is.na(abia$TaxiIn),] %>%
  group_by(Dest) %>%
  summarise(mean = mean(TaxiIn))

#7 Delays by Time Period for both arrivals and departures - AUS

period_dep_del = abia[((!is.na(abia$DepDelay))&(abia$Origin=="AUS")),] %>%
  group_by(Period_dep) %>%
  summarise(mean.del = mean(DepDelay))

period_arr_del = abia[((!is.na(abia$ArrDelay))&(abia$Dest=="AUS")),] %>%
  group_by(Period_arr) %>%
  summarise(mean.del = mean(ArrDelay))

## By DOW Level

period_dep_del_dow = abia[((!is.na(abia$DepDelay))&(abia$Origin=="AUS")),] %>%
  group_by(DayOfWeek,Period_dep) %>%
  summarise(mean.del = mean(DepDelay))

period_arr_del_dow = abia[((!is.na(abia$ArrDelay))&(abia$Dest=="AUS")),] %>%
  group_by(DayOfWeek,Period_arr) %>%
  summarise(mean.del = mean(ArrDelay))
```

1) What percentage of flights in and out of Austin are delayed ?

DEPARTURES
```{r echo=FALSE}
# BY MONTH
ggplot(depdel_month) + 
  geom_line(aes(x=Month, y=depdel.prt), color='blue') + 
  theme_bw(base_size=15) +
  scale_x_continuous(breaks = 1:12) +
  labs(title="% Departure Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay %",
       x = "Month")

```

```{r echo=FALSE,warning=FALSE}
# BY DAY OF MONTH
ggplot(depdel_dom) + 
  geom_line(aes(x=DayofMonth, y=depdel.prt), color='blue') + 
  theme_bw(base_size=15) +
  scale_x_continuous(breaks = 1:31) +
  labs(title="% Departure Delays from Austin - Day of Month", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay %",
       x = "Day of Month") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```

```{r echo=FALSE,warning=FALSE}
# BY DAY OF WEEK
ggplot(depdel_dow) + 
  geom_line(aes(x=DayOfWeek, y=depdel.prt), color='blue') + 
  theme_bw(base_size=15) +
  scale_x_continuous(breaks = 1:7) +
  labs(title="% Departure Delays from Austin - Day of Week", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay %",
       x = "Day of Week")
```


ARRIVALS
```{r echo=FALSE,warning=FALSE}
#BY MONTH
ggplot(arrdel_month) + 
  geom_line(aes(x=Month, y=arrdel.prt), color='red') + 
  theme_bw(base_size=15) +
  scale_x_continuous(breaks = 1:12) +
  labs(title="% Arrival Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay %",
       x = "Month")
```

```{r echo=FALSE}
# BY DAY OF MONTH
ggplot(arrdel_dom) + 
  geom_line(aes(x=DayofMonth, y=arrdel.prt), color='red') + 
  theme_bw(base_size=15) +
  scale_x_continuous(breaks = 1:31) +
  labs(title="% Arrival Delays from Austin - Day of Month", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay %",
       x = "Day of Month") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```

```{r echo=FALSE,warning=FALSE}
# BY DAY OF WEEK
ggplot(arrdel_dow) + 
  geom_line(aes(x=DayOfWeek, y=arrdel.prt), color='red') + 
  theme_bw(base_size=15) +
  scale_x_continuous(breaks = 1:7) +
  labs(title="% Departure Delays from Austin - Day of Week", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay %",
       x = "Day of Week")
```

2) What is the average amount of times by which flights in and out of Austin get delayed?

```{r echo=FALSE,warning=FALSE}
#BY MONTH
p1 = ggplot(mean_depdel_month) + 
  geom_line(aes(x=Month, y=mean.del), color='blue') + 
  theme_bw(base_size=10) +
  scale_x_continuous(breaks = 1:12) +
  labs(title="Mean Departure Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay Mean (mins)",
       x = "Month")

p2 = ggplot(mean_arrdel_month) + 
  geom_line(aes(x=Month, y=mean.del), color='red') + 
  theme_bw(base_size=10) +
  scale_x_continuous(breaks = 1:12) +
  labs(title="Mean Departure Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay Mean (mins)",
       x = "Month")

grid.arrange(p1,p2, nrow=1)

```

```{r echo=FALSE,warning=FALSE}
# BY DAY OF MONTH
p1 = ggplot(mean_depdel_dom) + 
  geom_line(aes(x=DayofMonth, y=mean.del), color='blue') + 
  theme_bw(base_size=10) +
  scale_x_continuous(breaks = 1:31) +
  labs(title="Mean Departure Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay Mean (mins)",
       x = "Day of Month") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))

p2 = ggplot(mean_arrdel_dom) + 
  geom_line(aes(x=DayofMonth, y=mean.del), color='red') + 
  theme_bw(base_size=10) +
  scale_x_continuous(breaks = 1:31) +
  labs(title="Mean Arrival Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay Mean (mins)",
       x = "Day of Month") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))

grid.arrange(p1,p2, nrow=1)
```

```{r echo=FALSE,warning=FALSE}
# BY DAY OF WEEK
p1 = ggplot(mean_depdel_dow) + 
  geom_line(aes(x=DayOfWeek, y=mean.del), color='blue') + 
  theme_bw(base_size=10) +
  scale_x_continuous(breaks = 1:31) +
  labs(title="Mean Departure Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay Mean (mins)",
       x = "Day of Week") 


p2 = ggplot(mean_arrdel_dow) + 
  geom_line(aes(x=DayOfWeek, y=mean.del), color='red') + 
  theme_bw(base_size=10) +
  scale_x_continuous(breaks = 1:31) +
  labs(title="Mean Arrival Delays from Austin", 
       caption="Source: ABIA dataset 2008",
       y="Flight Delay Mean (mins)",
       x = "Day of Week") 


grid.arrange(p1,p2, nrow=1)
```

3) What are the Origin(flights to Austin) and Destination(flights from Austin) with the most cancellation blues?
4) How are these split by the Cancellation Codes?
```{r echo=FALSE,warning=FALSE}
# BY ORIGIN
ggplot(orig_cancel, aes(x=Origin, y=tot.cancel)) + 
  geom_bar(stat='identity') +
  facet_wrap(~ CancellationCode, nrow = 3) +
  labs(title="Most Cancellations From ?", 
       caption="Source: ABIA dataset 2008",
       y="Total FLights Cancelled",
       x = "Origin Airport") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```
```{r echo=FALSE}

# BY DESTINATION
ggplot(dest_cancel, aes(x=Dest, y=tot.cancel)) + 
  geom_bar(stat='identity') +
  facet_wrap(~ CancellationCode, nrow = 3) +
  labs(title="Most Cancellations To ?", 
       caption="Source: ABIA dataset 2008",
       y="Total FLights Cancelled",
       x = "Destination Airport") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```

5) What are the airline carriers one should expect most delays with?

```{r echo=FALSE,warning=FALSE}
# CARRIER DELAYS
ggplot(na.omit(carrier_del), aes(x=UniqueCarrier, y=mean.del)) + 
  geom_bar(stat='identity') +
  labs(title="Which airline carriers to avoid?", 
       caption="Source: ABIA dataset 2008",
       y="Mean Delay Time",
       x = "Airline Carrier") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```

6) Which are the best Origin/Dest to expect pre-time arrival/departures to/from Austin ?

```{r echo=FALSE,warning=FALSE}
p1 = ggplot(pre_dep, aes(x=Origin , y=cnt.dep)) + 
  geom_bar(stat='identity') +
  labs(title="BEST airports to fly FROM, to Austin?", 
       caption="Source: ABIA dataset 2008",
       y="Total Pre Time Departures",
       x = "Airport") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))

p2 = ggplot(pre_arr, aes(x=Dest , y=cnt.arr)) + 
  geom_bar(stat='identity') +
  labs(title="BEST airports to fly TO, from Austin?", 
       caption="Source: ABIA dataset 2008",
       y="Total Pre Time Arrivals",
       x = "Airport") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))

grid.arrange(p1,p2, nrow=2)
```

7) Which airports do I expect to spend the maximum time Taxing In and Out?
```{r echo=FALSE,warning=FALSE}
p1 = ggplot(taxi_out, aes(x=Origin , y=mean)) + 
  geom_bar(stat='identity') +
  labs(title="Average Taxi Out times", 
       caption="Source: ABIA dataset 2008",
       y="Average time (mins)",
       x = "Airport") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))

p2 = ggplot(taxi_in, aes(x=Dest , y=mean)) + 
  geom_bar(stat='identity') +
  labs(title="Average Taxi In times", 
       caption="Source: ABIA dataset 2008",
       y="Average time (mins)",
       x = "Airport") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))

grid.arrange(p1,p2, nrow=2)
```

8) What are the times of the day I should expect the maximum arrival and departure delays in/out of Austin?
```{r echo=FALSE,warning=FALSE}
p1 = ggplot(period_dep_del, aes(x=Period_dep , y=mean.del)) + 
  geom_bar(stat='identity') +
  labs(title="Best time to fly out of Austin?", 
       caption="Source: ABIA dataset 2008",
       y="Average Dep Delay",
       x = "Time Of Day") 

p2 = ggplot(period_arr_del, aes(x=Period_arr , y=mean.del)) + 
  geom_bar(stat='identity') +
  labs(title="Best time to fly into Austin?", 
       caption="Source: ABIA dataset 2008",
       y="Average Arrival Delay",
       x = "Time Of Day") 

grid.arrange(p1,p2, nrow=1)
```

9) Are these trends better/worse on some days of the week? Whether I could beat this?
```{r echo=FALSE,warning=FALSE}
ggplot(period_dep_del_dow, aes(x=DayOfWeek, y=mean.del)) + 
  geom_bar(stat='identity') +
  facet_wrap(~ Period_dep, nrow = 7) +
  labs(title="Best time to fly out of Austin?", 
       caption="Source: ABIA dataset 2008",
       y="Average Dep Delay",
       x = "Day of Week") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```

```{r echo=FALSE,warning=FALSE}
ggplot(period_arr_del_dow, aes(x=DayOfWeek, y=mean.del)) + 
  geom_bar(stat='identity') +
  facet_wrap(~ Period_arr, nrow = 7) +
  labs(title="Best time to fly into Austin?", 
       caption="Source: ABIA dataset 2008",
       y="Average Arr Delay",
       x = "Day of Week") +
  theme(axis.text.x = element_text(angle=90, vjust=0.6))
```


#PORTFOLIO MODELLING

PORTFOLIO 1.
My Portfolio 1 consists of EFT's that have more of a global outlook like.

1. VT - global reach ETF
2. ACWX - has a global reach minus the United States
3. FXI - China focussed ETF
```{r echo=FALSE,warning=FALSE,message=FALSE}
#PORTFOLIO 1
mystocks = c("VT", "ACWX", "FXI")
myprices = getSymbols(mystocks, from = "2014-01-01")

#Adjusting for splits and dividends
for(ticker in mystocks) {
  expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
  eval(parse(text=expr))
}

#Combining all returns in a matrix
all_returns = cbind(ClCl(ACWXa),ClCl(FXIa),ClCl(VTa))
all_returns = as.matrix(na.omit(all_returns))


#Simulating different possible scenarios by bootstrap for PORTFOLIO 1

initial_wealth = 100000
sim1= foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.33, 0.33, 0.33) #
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
  }
  wealthtracker
}

# VAR Calculation
var = quantile(wealthtracker,c(0.05))
cat("The VAR for this portfolio is ", var)
```

PORTFOLIO 2.
My portfolio 2 is a safe retirement centric ETF portfolio. Long time hold with guaranteed yields.

1) VYM - retirement centric high dividend yield
2) VGLT - safe ETF with returns on long term
3) BOND - 3.4% safe yield return
4) PGX - 5.7% safe dividend returns, risk averse
```{r echo=FALSE,warning=FALSE}
#PORTFOLIO 2
mystocks2 = c("VYM", "VGLT", "BOND" , "PGX")
myprices2 = getSymbols(mystocks2, from = "2014-01-01")

#Adjusting for splits and dividends
for(ticker in mystocks2) {
  expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
  eval(parse(text=expr))
}

#Combining all returns in a matrix
all_returns2 = cbind(ClCl(VYMa),ClCl(VGLTa),ClCl(BONDa),ClCl(PGXa))
all_returns2 = as.matrix(na.omit(all_returns2))


#Simulating different possible scenarios by bootstrap for PORTFOLIO 2

initial_wealth = 100000
sim2= foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.25, 0.25, 0.25, 0.25) #
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns2, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
  }
  wealthtracker
}

#CALCULATING VAR
var = quantile(wealthtracker,c(0.05))
cat("The VAR for this portfolio is ", var)
```


PORTFOLIO 3.
My portfolio 3 is a diverse mix porfolio. High risk centric but higher returns on success too.

1) VGT - is a tech based ETF with some shares on Amazon too
2) XLY - is a big $12 Billion consumer fund
3) XBI - is a high risk ETF with money in a lot of medical startups
4) VNQ - is a real estate based ETF
```{r echo=FALSE,warning=FALSE}
#PORTFOLIO 3
mystocks3 = c("VGT", "XLY", "XBI" , "VNQ")
myprices3 = getSymbols(mystocks3, from = "2014-01-01")

#Adjusting for splits and dividends
for(ticker in mystocks3) {
  expr = paste0(ticker, "a = adjustOHLC(", ticker, ")")
  eval(parse(text=expr))
}

#Combining all returns in a matrix
all_returns3 = cbind(ClCl(VGTa),ClCl(XLYa),ClCl(XBIa),ClCl(VNQa))
all_returns3 = as.matrix(na.omit(all_returns3))


#Simulating different possible scenarios by bootstrap for PORTFOLIO 2

initial_wealth = 100000
sim3= foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.25, 0.25, 0.25, 0.25) #
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns3, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker[today] = total_wealth
  }
  wealthtracker
}

#CALCULATING VAR
var = quantile(wealthtracker,c(0.05))
cat("The VAR for this portfolio is ", var)
```
