---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r pressure}
library("ggplot2")
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra)
library(NbClust)
library(corrplot)
library(gridExtra)


getwd()
social <- read.csv("social_marketing.csv",header = TRUE)
#View(social)

social = read.csv("social_marketing.csv",row.names="X")
dim(social)
social = na.omit(social)
social_s = scale(social,center=TRUE, scale=TRUE)
social_n = social/rowSums(social)
social_ns = scale(social_n,center=TRUE, scale=TRUE)
#summary(social_ns)
cormat <- cor(social_ns)
corrplot(cormat,method = 'shade',type = 'upper')

# cormat <- round(cor(social_ns),2)
# cormat.topleft <- round(cor(social_ns[,1:18]),2)
# cormat.bottomright <- round(cor(social_ns[,19:36]),2)
# 
# 
# 
# #head(cormat)
# ggcorrplot(cormat)
# ggcorrplot(cormat.topleft)
# ggcorrplot(cormat.bottomright)

```


```{r pressure}
library(foreach)
set.seed(12)
#
# wss <- (nrow(social_ns)-1)*sum(apply(social_ns,2,var))
k_grid = seq(2, 7, by=1)
CH_grid = seq(2,7, by=1)
CH_grid
k_grid
N = nrow(social_ns)
# 
# for (i in 2:20) wss[i] <- sum(kmeans(social_ns, 
#    centers=i)$withinss)
# plot(1:20, wss, type="b", xlab="Number of Clusters",
#   ylab="Within groups sum of squares")
# 
# for (i in 2:20) {
#   cluster = kmeans(social_ns, i, nstart = 50)
#   B = cluster$Betweenss
#   W = cluster$tot.withinss
#   CH[i] = (B/W)*((N-i)/(i-1))
# } 
# 
# print(CH)
# 
# CH_grid = foreach(k = k_grid, combine='c') %do% {
# cluster_k = kmeans(social_ns, k, nstart=50) 
# W = cluster_k$tot.withinss
# B = cluster_k$betweenss
# CH = (B/W)*((N-k)/(k-1))
# CH
# }
# CH_grid
 for (i in 1:6) {
  cluster_k = kmeans(social_ns, k_grid[i], nstart=50)
  W = cluster_k$tot.withinss
  B = cluster_k$betweenss
  CH_grid[i] = (B/W)*((N-k_grid[i])/(k_grid[i]-1))
}
#CH_grid
#plot(k_grid,CH_grid)
ch_data = data.frame(k_grid, CH_grid)
chp = ggplot(data = ch_data) + 
  geom_point(mapping = aes(x = ch_data$k_grid, y = ch_data$CH_grid))


# Elbow method
wss = fviz_nbclust(social_ns, kmeans, method = "wss") +
  labs(subtitle = "Elbow method")

# Silhouette method
sil = fviz_nbclust(social_ns, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method")

grid.arrange(wss,sil, nrow = 1, ncol = 2)
grid.arrange(chp, nrow = 1, ncol = 1)

```
```{r pressure}
library(LICORS)

# K-Means Clustering with 3 clusters
set.seed(1)

# 3 clusters
set.seed(3)
k=3
clust3 = kmeans(social_ns, k, nstart=25)
print("cluster stat::k=3")
print("size of cluster")
clust3$size
print("withinss")
clust3$withinss
print("Total withinss")
clust3$tot.withinss
print("betweenss")
clust3$betweenss
print("CH")
(clust3$betweenss/clust3$tot.withinss)*((nrow(social_ns)-k)/(k-1))

# Visualizing the clusters on a dimensionally reduced plot
clus3plot = fviz_cluster(clust3, data = social_ns, 
             ellipse.type = "euclid", # Concentration ellipse
             ggtheme = theme_classic(),geom = c("point")
             )
# #Initializing kmeans++
# clust3p = kmeanspp(social_ns, k=k, nstart=50)
# clus3pplot = fviz_cluster(clust3p, data = social_ns, 
#              ellipse.type = "euclid", # Concentration ellipse
#              ggtheme = theme_classic(),geom = c("point")
#              )

# 4 clusters
set.seed(4)
k=4
clust4 = kmeans(social_ns, k, nstart=25)
print("cluster stat::k=4")
print("size of cluster")
clust4$size
print("withinss")
clust4$withinss
print("Total withinss")
clust4$tot.withinss
print("betweenss")
clust4$betweenss
print("CH")
(clust4$betweenss/clust4$tot.withinss)*((nrow(social_ns)-k)/(k-1))

# Visualizing the clusters on a dimensionally reduced plot
clus4plot = fviz_cluster(clust4, data = social_ns, 
             ellipse.type = "euclid", # Concentration ellipse
             ggtheme = theme_classic(),geom = c("point")
             )
#Initializing kmeans++
# clust4p = kmeanspp(social_ns, k=k, nstart=50)
# print("cluster stat::")
# print("size of cluster")
# clust4p$size
# print("withinss")
# clust4p$withinss
# print("Total withinss")
# clust4p$tot.withinss
# print("betweenss")
# clust4p$betweenss
# print("CH")
# (clust4p$betweenss/clust4p$tot.withinss)*((nrow(social_ns)-k)/(k-1))
# 
# clus3pplot = fviz_cluster(clust4p, data = social_ns, 
#              ellipse.type = "euclid", # Concentration ellipse
#              ggtheme = theme_classic(),geom = c("point")
#              )
# grid.arrange(clus3plot,clus3pplot,ncol = 2)
# 

# 5 clusters
set.seed(3)
k=5
clust5 = kmeans(social_ns, k, nstart=25)
print("cluster stat::k=5")
print("size of cluster")
clust5$size
print("withinss")
clust5$withinss
print("Total withinss")
clust5$tot.withinss
print("betweenss")
clust5$betweenss
print("CH")
(clust5$betweenss/clust5$tot.withinss)*((nrow(social_ns)-k)/(k-1))

# Visualizing the clusters on a dimensionally reduced plot
clus5plot = fviz_cluster(clust5, data = social_ns, 
             ellipse.type = "euclid", # Concentration ellipse
             ggtheme = theme_classic(),geom = c("point")
             )
#Initializing kmeans++
# clust5p = kmeanspp(social_ns, k=k, nstart=50)
# clus3pplot = fviz_cluster(clust5p, data = social_ns, 
#              ellipse.type = "euclid", # Concentration ellipse
#              ggtheme = theme_classic(),geom = c("point")
#              )
# grid.arrange(clus3plot,clus3pplot,ncol = 2)

# 6 clusters
set.seed(3)
k=6
clust6 = kmeans(social_ns, k, nstart=25)
print("cluster stat::")
print("size of cluster")
clust6$size
print("withinss")
clust6$withinss
print("Total withinss")
clust6$tot.withinss
print("betweenss")
clust6$betweenss
print("CH")
(clust6$betweenss/clust6$tot.withinss)*((nrow(social_ns)-k)/(k-1))

# Visualizing the clusters on a dimensionally reduced plot
clus6plot = fviz_cluster(clust6, data = social_ns, 
             ellipse.type = "euclid", # Concentration ellipse
             ggtheme = theme_classic(),geom = c("point")
             )

grid.arrange(clus3plot,clus4plot,clus5plot,clus6plot, nrow = 2, ncol = 2)


#perform PCA

pr.out = prcomp(social_ns,scale=TRUE)
plot(pr.out)
#summary(pr.out)

scores = pr.out$x
qplot(scores[,1], scores[,2], xlab='Component 1', colour = clust2$cluster, ylab='Component 2')

loadings = pr.out$rotation
o1 = order(loadings[,1], decreasing=TRUE)
print("PC1 head")
colnames(social_ns)[head(o1,10)]
print("PC1 tail")
colnames(social_ns)[tail(o1,10)]
 
o2 = order(loadings[,2], decreasing=TRUE)
print("PC2 head")
colnames(social_ns)[head(o2,10)]
print("PC2 tail")
colnames(social_ns)[tail(o2,10)]


```
```{r cluster level summaries}
library(mosaic)
library(tidyverse)

attach(social)
social_grp2 = social
social_grp2$influencers = photo_sharing+chatter+sports_fandom+travel+outdoors
social_grp2$health.enthusiasts= food+cooking+home_and_garden+health_nutrition+personal_fitness+sports_playing
social_grp2$elderly = politics+news+religion+current_events+family+parenting+shopping+business+small_business+eco+automotive
social_grp2$teens = tv_film+music+online_gaming+beauty+dating+fashion+college_uni+school+art+crafts+computers
social_grp2$others = spam+uncategorized+adult
# append cluster assignment
mydata <- data.frame(social_grp2, fit$cluster)

# get cluster means 
agg = aggregate(social,by=list(clust2$cluster),FUN=mean)
agg_T = t(agg)

#agg_T %>% mutate(stDev = apply(agg_T,1,sd))
agg_T
#plot(agg_T)
# append cluster assignment
mydata <- data.frame(social_grp2, fit$cluster)
mydata <- data.frame(mydata, clust2$cluster)
#colnames(mydata)

mydata_summ = mydata %>%
  group_by(clust2.cluster)  %>%  # group the data points by model name
  summarize(health.enthusiasts.mean = mean(health.enthusiasts),
            influencers.mean = mean(influencers),
            elderly.mean = mean(elderly),
            teens.mean = mean(teens),
            others.mean = mean(others))  # calculate a mean for #each model

mydata_summ
# # still not in order...
# ggplot(mydata_summ, aes(x=cluster, y=health.enthusiasts.mean)) + 
#   geom_bar(stat='identity') + 
#   coord_flip()
# 
# # reorder the x labels
# ggplot(mydata_summ, aes(x=reorder(fit.cluster, health.enthusiasts.mean), y=health.enthusiasts.mean)) + 
#   geom_bar(stat='identity') + 
#   coord_flip()

par(mfrow=c(2,2))
g = ggplot(mydata, aes(x=elderly))
g + geom_density(aes(fill=factor(clust2.cluster)))

g = ggplot(mydata, aes(x=health.enthusiasts))
g + geom_density(aes(fill=factor(clust2.cluster)))

g = ggplot(mydata, aes(x=teens))
g + geom_density(aes(fill=factor(clust2.cluster)))

g = ggplot(mydata, aes(x=others))
g + geom_density(aes(fill=factor(clust2.cluster)))




```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
